### 2017-02-08 doc/manuscript.Rmd
Expanded on wording of PLS methods as per Tony's email

# Supplemental Methods

@@

Generalized estimating equations (GEE) is a technique similar to mixed effects
modeling, except it calculates the marginal population estimates compared to the
subject-specific estimates in mixed effects models. GEE is well suited and commonly
used in longitudinal cohort studies, especially given its capacity to handle
missed visits.

@@

For the confounders, they were chosen based on previous literature, from
directed acyclic graph (DAG) [@Greenland1999a] recommendations, and from
QIC. The DAG recommendations were obtained from using the DAGitty software
[@Textor2011a], http://dagitty.net/. See `r cite_sf('dag_is')` and 
`r cite_sf('dag_bcf')`) for the DAG model. See `r cite_st('qic')` for the
comparison of various models with different covariates using QIC. While the
final GEE model selected as best fitting differed between insulin sensitivity
and beta-cell function measures, the differences in the QIC values were less than 10
between many of the models, suggesting similar fit. As such, we selected the
model that had the fewest covariates and that had similar fit between the
outcome measures. {{NEFA comment for inclusion here}}

@@

<!--
The number of components to extract from PLS was determined by 10-fold CV.
Dataset was split into a training set and a testing set to determine the
predictability of the given PLS model. Final results presented use the full dataset.a
-->

We used internal 10-fold cross-validation to determine which components to
extract from the PLS analysis. Based on the internal cross-validation, the first
two components gave the highest amount of explained variance (data not shown),
of which we decided to use these two components in the final results.

The predictive capability of the PLS models were tested using cross-validation. 
The data set was randomly split in half into a training set and a testing set. 
After specifying the model on the training set, results were compared using the 
testing set to determine whether how predictive the model was given a new
dataset. Final results shown in the figures use the full dataset (rather than a
training or testing set)

No prevalent diabetes cases were included in the PLS
analysis.

>>>>

# Supplemental Material

@@

Generalized estimating equations (GEE) is a technique similar to mixed effects
modeling, except it calculates the marginal population estimates compared to the
subject-specific estimates in mixed effects models. GEE is well suited to and commonly
used in longitudinal cohort studies, especially given its capacity to handle
missed visits.

@@

For the confounders, they were chosen based on previous literature, from
directed acyclic graph (DAG) [@Greenland1999a] recommendations, and from
QIC. The DAG recommendations were obtained from using the DAGitty software
[@Textor2011a], http://dagitty.net/. See `r cite_sf('dag_is')` and 
`r cite_sf('dag_bcf')`) for the DAG model. See `r cite_st('qic')` for the
comparison of various models with different covariates using QIC. While the
final GEE model selected as best fitting differed between insulin sensitivity
and beta-cell function measures, the differences in the QIC values were less than 10
between many of the models, suggesting similar fit. As such, we selected the
model that had the fewest covariates and that had similar fit between the
outcome measures. In particular for NEFA, it is used substantially as a source
of fatty acids in hepatic TAG production [@Barrows2006a] and may be a strong
confounder.

@@

PLS is a technique used for multivariate, high dimensionality datasets where
there is a potential or likely underlying structure to the data. Because it uses
a response variable (i.e. the outcome or y variable) when identifying the 
underlying structure it is known as a supervised statistical method, which gives
it greater predictive power when using the model against updated or new data.
PLS generates a number of components based on the number of variables provided.
We used internal 10-fold cross-validation to determine which components to
extract from the PLS analysis. Based on the internal cross-validation, the first
two components gave the highest amount of explained variance (data not shown),
of which we decided to use these two components in the final results.

The predictive capability of the PLS models were tested using cross-validation. 
The data set was randomly split in half into a training set and a testing set. 
After specifying the model on the training set, results were compared using the 
testing set to determine whether how predictive the model was given a new
dataset. Final results shown in the figures use the full dataset (rather than a
training or testing set). No prevalent diabetes cases were included in the PLS 
analysis.
